Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-12-16 manuscript/manuscript.tex
Revised chart margin and height/width sizing to eliminate blank space
Incorporating JY's suggestions on adding parameters to chart sizing, including height, width, and margins (top, left, bottom, right). Removed the "-crop" PDF files and replaced with new charts generated by the R programs.

\begin{figure}[tbp]
	\centering
	\includegraphics[width = \textwidth]{SRs_by_Agency-crop.pdf}
  	\caption{311 SR counts by agency for the time period 2022--2023.}
	\label{fig:SRcountbyAgency}
\end{figure}

@@

\begin{figure}[tbp]
 \centering
  \includegraphics[width = \textwidth]{SR_by_Complaint_Type-crop.pdf} 
  \caption{Top 20 complaint types for the 311 SR during 2022--2023.} 
  \label{fig:SR_complaints}
\end{figure}

@@

The \textit{Mostly Empty} category includes fields such as
\texttt{taxi\_company\_borough}, \texttt{due\_date}
\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.

Of particular note is that several of the \textit{Mostly Empty} fields
are used by only one or two Agencies. Example:

The \textit{Partially Empty} category includes fields such as
\texttt{location\_type}, \texttt{landmark}, 
and \texttt{cross\_street\_1 \& \_2}.

And the \textit{Few/None Empty} 
category includes \texttt{created\_date}, \texttt{complaint\_type},
\texttt{agency}, and \texttt{status}.

For some analysis efforts, it may be prudent to inquire 
further as to why some fields are mostly or almost 
always blank. Figure~\ref{fig:blank_fields} is 
a graphic depiction of total empty (blank \& N/As) for each 
field illustrating the grouping into the Most, 
Partial, and Few/None categories.

\begin{figure}[tbp]
	\centering
  	\includegraphics[width=\textwidth]{BlankFields-crop.pdf}
	\caption{Number and percentage of empty/blank entries for the
          311 SR data during 2022--2023.}
	\label{fig:blank_fields}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  % First figure
  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
    \centering
    \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
    \caption{Distribution of negative response times.}
    \label{fig:negative-duration-violin}
  \end{subfigure}
  \hfill % Horizontal space between subfigures
  % Second figure
  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
    \centering
    \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
    \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days}
    \label{fig:resolution-violin}
  \end{subfigure}
  % Main figure caption
  \caption{Violin plots for negative response times and post-closed
    resolution during 2022--2023.}
  \label{fig:violin-plots}
\end{figure}

@@

\jy{The font sizes of annotations in Figure 4 and 5 need to be
  increased. In addition, the size of the pdf figure could be made
  arguments to the R function that generated them (instead of being
  fixed in within the function).}

@@

\jy{Could refer to cross-platform data format such as Arrow
  \citep{bates2024csv}.}

>>>>

\begin{figure}[tbp]
	\centering
	\includegraphics[width = \textwidth]{SRs_by_Agency.pdf}
  	\caption{311 SR counts by agency for the time period 2022--2023.}
	\label{fig:SRcountbyAgency}
\end{figure}

@@

\begin{figure}[tbp]
 \centering
  \includegraphics[width = \textwidth]{SR_by_Complaint_Type.pdf} 
  \caption{Top 20 complaint types for the 311 SR during 2022--2023.} 
  \label{fig:SR_complaints}
\end{figure}

@@

The Mostly Empty category includes:
\texttt{taxi\_company\_borough}, \texttt{due\_date}
\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}. The 
Partially Empty category includes:
\texttt{location\_type}, \texttt{landmark}, 
and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty 
category includes \texttt{created\_date}, \texttt{complaint\_type},
\texttt{agency}, and \texttt{status}. Of particular note is that 
several of the Mostly Empty fields are used by only one or 
two Agencies (See: \autoref{tab:field-usage-summary}).

\begin{table}[htbp]
    \centering
    \caption{Fields used by only one or two Agencies with limited non-blank entires}
    \label{tab:field-usage-summary}
    \begin{tabular}{l r r p{4cm}}
        \toprule
        \textbf{Field} & \textbf{\# of entries} & \textbf{\% non-blank} & \textbf{Agencies using} \\
        \midrule
        \texttt{taxi\_company\_borough} & 3,567  & 0.06\% & TLC \\
        \texttt{vehicle\_type}           & 18,642 & 0.29\% & TLC, NYPD \\
        \texttt{due\_date}               & 24,414 & 0.38\% & DSNY \\
        \texttt{taxi\_pick\_up\_location} & 64,954 & 1.02\% & TLC \\
        \bottomrule
    \end{tabular}
\end{table}

For some analysis efforts, it may be prudent to inquire 
further as to why some fields are mostly or almost 
always blank. Figure~\ref{fig:blank_fields} is 
a graphic depiction of total empty (blank \& N/As) for each 
field illustrating the grouping into the Mostly Empty, 
Partially Empty, and Few/None Empty categories.

\begin{figure}[tbp]
	\centering
  	\includegraphics[width=\textwidth]{BlankFields.pdf}
	\caption{Number and percentage of empty/blank entries for the
          311 SR data during 2022--2023.}
	\label{fig:blank_fields}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  % First figure
  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
    \centering
    \includegraphics[width=\textwidth]{negative_duration_SR_violin.pdf}
    \caption{Distribution of SR durations.}
    \label{fig:negative-duration-violin}
  \end{subfigure}
  \hfill % Horizontal space between subfigures
  % Second figure
  \begin{subfigure}[t]{0.495\textwidth} % Adjust width as needed
    \centering
    \includegraphics[width=\textwidth]{post_closed_violin_chart.pdf}
    \caption{Post-closed \texttt{resolution\_action\_update\_date}(s) >30 days}
    \label{fig:resolution-violin}
  \end{subfigure}
  % Main figure caption
  \caption{Violin plots for negative SR durations and post-closed
    resolution updates (2022--2023.}
  \label{fig:violin-plots}
\end{figure}

@@

@@

One promising alternative to reducing data storage requirements 
for this large 311 SR dataset (3.7GB) would be employment of 
the cross-platform Apache Arrow package. Apache Arrow (arrow.apache.org) 
is an innovative, cross-language development platform designed for 
in-memory data. It defines a standard for representing columnar data, 
enabling efficient analytics and real-time data processing. As such,
it offers dramatic improvements in file storage size, as well as 
efficient analysis and cross-platform support for R, Python, 
and Julia programming languages. Remarkable storage size 
reduction (within several orders of magnitude) is possible. \citep{bates2024csv}.
